Linear instability underlying synchronous waving of marine grass in flow

### 2014-09-21 Paper/grass.tex
minor change

Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in
the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is shown in the inset in figure \ref{Re_vs_delta}b. 
The observed frequencies are associated with the peaks in velocity spectra, frequency of $\monami$ and frequency of 
vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies. 
Regardless of the assumptions, the introduction of vegetation drag is observed to damp the linear modes\cite{Delangre06}.

@@

Similar phenomenon of large amplitude coherent oscillation of terrestrial canopies in atmospheric flow known as \textit{honami}\cite{Inoue56,Raupach96}, have also been observed.
A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically\cite{Vivoni98,Nepf00}. 
Another major difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid, whereas aquatic canopies are buoyant\cite{Vivoni98,Ghisal02}. 
Despite these differences, in the framework of our model, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represents the hydrodynamic instability for the terrestrial case with stiff grass blades. We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large grass density, so as to make Mode 2 irrelevant. We recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case.

The deviation of our model predictions from the observed may be attributed to the various simplification we have made in our model. 
In a real canopy, the drag coefficients are known to vary from bottom to tip of the grass blades\cite{Vivoni98,Nepf00}. 
The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy\cite{Ghisal02}. 
Although an improved analysis of flow with these variation in the model might lead to a better agreement between the observed and the predicted quantities, these variation are not central to the mechanism leading to the instability of flow.

In conclusion, we note that the threshold flow condition observed in the field and in lab experiments arises due to the presence of vegetation drag. 
We show the origin of this threshold using a consistent continuum hydrodynamic stabiity analysis for the shear flow that develops as a result of the drag. 
Such an analysis not only explains quantitative observations, but also provides an understanding of the hydrodynamic instability mechanism. 
The extent of the structure of Mode 2 throughout the depth of the water column implies an efficient transport of suspended tracers compared to that due to the localized structure of Mode 1 near the canopy top. 
Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, etc. and therefore is expected to have a wider impact.

>>>>

Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in
the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is shown in the inset in figure \ref{Re_vs_delta}b. 
The observed frequencies are associated with the peaks in velocity spectra, frequency of $\monami$ and frequency of 
vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies. 
%Regardless of the assumptions, the introduction of vegetation drag is observed to damp the linear modes\cite{Delangre06}.

@@

Similar phenomenon of large amplitude coherent oscillation of terrestrial canopies in atmospheric flow known as \textit{honami}\cite{Inoue56,Raupach96}, have also been observed.
A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically\cite{Vivoni98,Nepf00}. 
Another major difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid, whereas aquatic canopies are buoyant\cite{Vivoni98,Ghisal02}. 
Despite these differences, in the framework of our model, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represents the hydrodynamic instability for the terrestrial case with stiff grass blades. We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large grass density, so as to make Mode 2 irrelevant. We recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case. 
\newline
The deviation of our model predictions from the observed may be attributed to the various simplification we have made in our model. 
In a real canopy, the drag coefficients are known to vary from bottom to tip of the grass blades\cite{Vivoni98,Nepf00}. 
The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy\cite{Ghisal02}. 
Although an improved analysis of flow with these variation in the model might lead to a better agreement between the observed and the predicted quantities, these variation are not central to the mechanism leading to the instability of flow.
\newline
In conclusion, we note that the threshold flow condition observed in the field and in lab experiments arises due to the presence of vegetation drag. 
We show the origin of this threshold using a consistent continuum hydrodynamic stabiity analysis for the shear flow that develops as a result of the drag. 
Such an analysis not only explains quantitative observations, but also provides an understanding of the hydrodynamic instability mechanism. 
The extent of the structure of Mode 2 throughout the depth of the water column implies an efficient transport of suspended tracers compared to that due to the localized structure of Mode 1 near the canopy top. 
Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, etc. and therefore is expected to have a wider impact.
